---
title: "ASOFAMECH"
format: pdf
code-annotations: hover
---

## Metodology and results.

### Statistical Analysis.

Statistical analysis was performed with R (Version 4.2.3) and Rstudio (Version 2023.03.0+386) statisticals softwares. All data were analyzed using one-tailed tests with a significance level of α = 0.05. Confidence intervals (CI) were constructed with a confidence level of 1 − α = 0.95, using the t Student distribution.

### Intra class correlation for evaluators.

Intra-Class Correlation (ICC) is a measure used in statistics to determine how strongly units in the same group resemble each other. It's often used in reliability studies where you want to assess the consistency or homogeneity within groups.

Negative ICC estimates are possible and can be interpreted as indicating that the true ICC is low, that is, two members chosen randomly from any class vary almost as much as any two randomly chosen members of the whole population.

```{r include=FALSE}
library(knitr)
library(DT)
library(readxl)
library(irr)
library(tidyverse)
library(kableExtra)

opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)
```

```{r}
DATA<- read_excel("Tabulación Resultados Amaru.xlsx", 
    sheet = "Amaru")

DATA[,5:516] <- DATA[,5:516]  %>% 
  mutate(across(everything(), ~ifelse(. == "No aplica", NA_character_, as.character(.)))) %>% 
  mutate(across(where(is.character), as.numeric))

calculate_icc <- function(data, group_col, cols) {
  data %>%
    select(group_col, all_of(cols)) %>% 
    group_by(!!sym(group_col)) %>% 
    nest() %>%
    mutate(
      icc_results = map(data, ~icc(select(.x, where(~any(!is.na(.)))))),
      icc_value = map_dbl(icc_results, ~.x$value),
      p_value = map_dbl(icc_results, ~.x$p.value)
    )
}

column_ranges <- list(
  item1_time1 = c(5:20),
  item1_time2 = c(21:36),
  item2_time1 = c(37:52),
  item2_time2 = c(53:68),
  item3_time1 = c(69:84),
  item3_time2 = c(85:100),
  item4_time1 = c(101:116),
  item4_time2 = c(117:132),
  item5_time1 = c(133:148),
  item5_time2 = c(149:164),
  item6_time1 = c(165:180),
  item6_time2 = c(181:196),
  item7_time1 = c(197:212),
  item7_time2 = c(213:228),
  item8_time1 = c(229:244),
  item8_time2 = c(245:260),
  item9_time1 = c(261:276),
  item9_time2 = c(277:292),
  item10_time1 = c(293:308),
  item10_time2 = c(309:324),
  item11_time1 = c(325:340),
  item11_time2 = c(341:356),
  item12_time1 = c(357:372),
  item12_time2 = c(373:388),
  item13_time1 = c(389:404),
  item13_time2 = c(405:420),
  item14_time1 = c(421:436),
  item14_time2 = c(437:452),
  item15_time1 = c(453:468),
  item15_time2 = c(469:484),
  item16_time1 = c(485:500),
  item16_time2 = c(501:516)
)


results_list <- lapply(column_ranges, function(cols) {
  calculate_icc(DATA, "Grupo", cols)
})

names(results_list) <- names(column_ranges)

results_list <- lapply(names(column_ranges), function(name) {
  icc_results <- calculate_icc(DATA, "Grupo", column_ranges[[name]])
  # Adding item_time column for identification
  icc_results <- icc_results %>%
    mutate(item_time = name) %>%
    select(-data, -icc_results) %>% 
    pivot_longer(cols = c(icc_value, p_value), names_to = "measure", values_to = "value") %>%
    mutate(variable = paste(item_time, measure, sep = "_")) %>%
    select(-item_time, -measure) %>%
    pivot_wider(names_from = variable, values_from = value)
  
  return(icc_results)
})

# Combine all results into a single dataframe
final_results <- bind_rows(results_list) %>%
  group_by(Grupo) %>%
  summarise(across(everything(), ~mean(., na.rm = TRUE)))
final_results <- final_results %>% mutate_all(~round(., 2))
kable(final_results[,1:5], format = "latex", table.attr = "class='display'") %>%
  kable_styling(font_size = 7)
```

### Mean comparison between pre- and post-intervention,

For the mean comparison between pre- and post-interventions for each item score and total item scores, the Wilcoxon signed-rank test was used, which is a non-parametric alternative to the paired t-test. This test was used when the assumptions of the paired t-test (such as the normal distribution of differences) were not met. It is appropriate to compare two related samples when the data are ordinal or when interval-level measurements do not follow a normal distribution.

#### Scores for each item.

```{r fig.dpi=1000, fig.height=8, fig.width= 10}
# Function to calculate row means for given columns
calculate_row_means <- function(df, cols) {
  rowMeans(df[, cols, drop = FALSE], na.rm = TRUE)
}

# Calculating row means for each item_time and storing in a new dataframe
mean_values <- map(column_ranges, ~calculate_row_means(DATA, .x))

# Binding the results into a single dataframe
mean_values_df <- bind_cols(mean_values)
names(mean_values_df) <- paste("mean", names(column_ranges), sep = "_")

reshaped_df <- mean_values_df %>%
  pivot_longer(everything(), names_to = "item_time", values_to = "mean") %>%
  mutate(
    item = gsub("_time[12]", "", item_time),
    time = gsub(".*_(time[12])", "\\1", item_time)
  ) %>%
  select(-item_time) %>%
  pivot_wider(names_from = time, values_from = mean)


results_df <- reshaped_df %>%
  rowwise() %>%
  mutate(
    mean_time1 = mean(time1, na.rm = TRUE),
    mean_time2 = mean(time2, na.rm = TRUE),
    ci_lower_time1 = mean_time1 - qt(0.975, df=16-1) * sd(time1, na.rm = TRUE) / sqrt(16),
    ci_upper_time1 = mean_time1 + qt(0.975, df=16-1) * sd(time1, na.rm = TRUE) / sqrt(16),
    ci_lower_time2 = mean_time2 - qt(0.975, df=16-1) * sd(time2, na.rm = TRUE) / sqrt(16),
    ci_upper_time2 = mean_time2 + qt(0.975, df=16-1) * sd(time2, na.rm = TRUE) / sqrt(16),
    wilcoxon_p_value = wilcox.test(time1, time2, paired = TRUE)$p.value
  ) %>%
  select(item, starts_with("mean"), starts_with("ci"), wilcoxon_p_value)


graph_data <- results_df %>%
  pivot_longer(
    cols = c(mean_time1, mean_time2),
    names_to = "time",
    values_to = "mean"
  ) %>%
  mutate(
    time = factor(time, levels = c("mean_time1", "mean_time2")),
    ci_lower = if_else(time == "mean_time1", ci_lower_time1, ci_lower_time2),
    ci_upper = if_else(time == "mean_time1", ci_upper_time1, ci_upper_time2)
  )

graph_data <- graph_data %>%
  mutate(
    item = as.numeric(str_replace(item, "mean_item", "")),
    time = as.numeric(str_replace(time, "mean_time", "")),
    time = factor(time, levels = c(1, 2), labels = c("Pre intervención", "Post intervención"))
  ) %>% select(1,7,8,9,10,6)

ggplot(graph_data, aes(x = as.factor(item), y = mean, fill = as.factor(time))) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7, alpha=0.7) +
  geom_errorbar(
    aes(ymin = ci_lower, ymax = ci_upper), 
    width = 0.2, 
    position = position_dodge(width = 0.8)
  ) +
  geom_text(
    data = subset(graph_data, time == "Post intervención"),
    aes(label = sprintf("p = %.2f", wilcoxon_p_value), y = ci_upper + 0.05),
    position = position_dodge(width = 0.8),
    vjust = 0,
    size = 3
  ) +
  scale_fill_manual(values = c("Pre intervención" = "skyblue", "Post intervención" = "blue")) +
  labs(x = "Item", y = "Media del puntaje (IC. 95%)") +
  theme_minimal() +
  theme(legend.title = element_blank())
```

#### Total item scores.

```{r fig.dpi=1000, fig.height=8, fig.width= 4, out.width='40%'}

# Assuming mean_values_df has columns like item1_time1, item1_time2, ..., item16_time1, item16_time2

# Calculate row sums for time1 across all items
mean_values_df$row_sum_time1 <- rowSums(mean_values_df[grepl("time1$", names(mean_values_df))], na.rm = TRUE)

# Calculate row sums for time2 across all items
mean_values_df$row_sum_time2 <- rowSums(mean_values_df[grepl("time2$", names(mean_values_df))], na.rm = TRUE)

# Calculate the mean
mean_time1 <- mean(mean_values_df$row_sum_time1, na.rm = TRUE)
mean_time2 <- mean(mean_values_df$row_sum_time2, na.rm = TRUE)

# Calculate t-student confidence intervals
se_time1 <- sd(mean_values_df$row_sum_time1, na.rm = TRUE) / sqrt(nrow(mean_values_df))
se_time2 <- sd(mean_values_df$row_sum_time2, na.rm = TRUE) / sqrt(nrow(mean_values_df))
ci_lower_time1 <- mean_time1 - qt(0.975, df=nrow(mean_values_df)-1) * se_time1
ci_upper_time1 <- mean_time1 + qt(0.975, df=nrow(mean_values_df)-1) * se_time1
ci_lower_time2 <- mean_time2 - qt(0.975, df=nrow(mean_values_df)-1) * se_time2
ci_upper_time2 <- mean_time2 + qt(0.975, df=nrow(mean_values_df)-1) * se_time2

# Perform Wilcoxon signed-rank test
wilcoxon_test <- wilcox.test(mean_values_df$row_sum_time1, mean_values_df$row_sum_time2, paired = TRUE)
p_value <- wilcoxon_test$p.value

# Create a new dataframe with the results
results_df_2 <- data.frame(
  time = c(1, 2),
  mean = c(mean_time1, mean_time2),
  ci_lower = c(ci_lower_time1, ci_lower_time2),
  ci_upper = c(ci_upper_time1, ci_upper_time2),
  p_value=p_value
)
results_df_2$time = factor(results_df_2$time, levels = c(1, 2), labels = c("Pre intervención", "Post intervención"))

max_y <- max(results_df_2$ci_upper, na.rm = TRUE)

ggplot(results_df_2, aes(x = time, y = mean, fill = time)) +
  geom_bar(stat = "identity", position = position_dodge(), width = 0.9, alpha = 0.7) +
  geom_errorbar(
    aes(ymin = ci_lower, ymax = ci_upper),
    width = 0.2,
    position = position_dodge(0.7)
  ) + 
  geom_text(
    aes(x = 1.5, y = max_y, label = sprintf("p = %.2f", results_df_2$p_value[1])),
    vjust = -0.5,
    size = 4
  ) +
  scale_fill_manual(values = c("Pre intervención" = "skyblue", "Post intervención" = "blue")) +
  labs(x = "", y = "Media del puntaje total de todos los items (IC. 95%)") +
  theme_minimal() +
  theme(legend.position = "none")
```
